---
title: "Macrodemography analyses"
author:
 - Jacob Socolar
 - Adriaan Dokter
date: "`r Sys.Date()`"
output: html_document
params:
  erd_path: "~/Dropbox/macrodemography/erd/erd.db"
  output_path: "~/Dropbox/macrodemography_refactor/data/residents"
  years: !r c(2006:2019)
  extent_space: !r
    data.frame(
      min_lon=-125, 
      max_lon=-66, 
      min_lat=24, 
      max_lat=50
    )
  period: !r c("spring", "fall")
  time_grid: 7 # resolution of the time grid
  tgrid_min: !r c(13, 40)
  tgrid_max: !r c(16, 43)
  # we need to constrain altitudes, particularly in the western US, where 
  # high altitudes are snowbound and inaccessible in spring. We don't want huge
  # variation in the elevations visited between April and October, especially
  # because that variation will probably depend on winter snow conditions, 
  # leading to interannual idosyncracies that are noisy and possibly correlated
  # with weather variables of interest. The appropriate altitudinal cutoff will
  # be higher in the Sierras and southern Rockies than in the Cascades and
  # northern Rockies.
  max_altitude: 2000
  max_altitude_above_lat42: 1500
  effort_thresholds: !r
    data.frame(
      dist_max=3, 
      time_min=5/60, 
      time_max=1, 
      cci_min=0
    )
  hexagon_area_large: 70000 # desired approx. area of large cells (sq km)
  hexagon_area_small: 300 # desired approx. area of small cells
  n_small_min: 10 # minimum number of small cells to compute abundance index for large cell
  
  daymet: !r
    data.frame(
      label=c("tmax_summer","tmax_winter","swe"), 
      variable=c("tmax","tmax","swe"),
      date_min=c("01-01","07-01","12-01"),
      date_max=c("02-28","08-31","03-15")
    )
  species_to_process: !r c("carwre")
  always_import_checklists: FALSE
  always_filter_checklists: FALSE
  always_resample_bootstrap: FALSE
  always_run_variance_test: FALSE
  quiet: TRUE
  region: "eastern_us"
---

```{r packages, include=FALSE, echo=FALSE}
# Install an older version of dggridR from source. See:
# https://github.com/r-barnes/dggridR/issues/63#issuecomment-1454929653
remotes::install_github("r-barnes/dggridR", ref = "ec2a040")

# we make sure we have the same version of the erdPackage as this .rmd
package_path <- strsplit(rstudioapi::getSourceEditorContext()$path, "inst")[[1]][1]
install.packages(package_path, repos = NULL, type = 'source')

# package load
library(macrodemography)
library(data.table) # don't worry about openMP warnings on Mac.
library(brms)
library(ggplot2)
library(dplyr)
library(magrittr)
library(sf)
library(dtplyr) # enable dplyr for data.table
library(assertthat)
library(ebirdst)
library(fasterize)
library(dggridR)
library(rgee)

# this package also needs a working version of rcmdstan:
# see https://mc-stan.org/cmdstanr/
cmdstanr::check_cmdstan_toolchain()
assert_that(cmdstanr::cmdstan_version() >= "2.31")
```

```{r misc-setup}
# set color scales
cols_bd <- c(hsv(seq(0,.17,length.out = 100),1,seq(.9,.6,length.out = 100)), hsv(seq(.45,.65, length.out = 100),1,seq(.6,1,length.out = 100)))
cols_bd2 <- c(hsv(seq(0,.17,length.out = 100),seq(1, .2, length.out = 100),.9), hsv(seq(.45,.65, length.out = 100),seq(.2, 1, length.out = 100),.9))

# blank ggplot2 theme
blank_theme <-
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank()
  )

western_states <- c("arizona", "california", "colorado", "idaho", "montana",
                    "nevada", "new mexico", "oregon", "utah", "washington",
                    "wyoming")

if(params$region == "eastern_us"){
  region_of_interest <- spData::us_states %>% filter(!(tolower(NAME) %in% western_states))
} else if(params$region == "western_us") {
  region_of_interest <- spData::us_states %>% filter(toupper(NAME) %in% western_states)
} else if(params$region == "conus") {  
  region_of_interest <- spData::us_states
} else if(params$region == "north_america") {
  region_of_interest <- ne_states(country =  c("United States of America","Canada"), returnclass = "sf") %>%
  filter(region != "Northern Canada") %>% 
  filter(!name %in% c("Hawaii", "Alaska"))
} else{
  region_of_interest <- spData::us_states %>% filter(tolower(NAME) %in% params$region)
}
# NOTE: this adds additional cropping of the region of interest
# adjust if necessary
raster_of_interest <- fasterize::fasterize(region_of_interest,
                                   raster::raster(ncol=1000, nrow = 1000, 
                                                  xmn = -125, xmx = -66, 
                                                  ymn =24, ymx = 50))
```


```{r define-time-extent}
# we can't define this in the run params because years isn't yet defined
extent_time <-
  data.frame(
    period = params$period,
    tgrid_min = params$tgrid_min, 
    tgrid_max = params$tgrid_max, 
    year_min = min(params$years), 
    year_max = max(params$years)
  )

```


```{r create=dggrids}
# for an area of ~ 70000 km^2, we get a resolution of 6:
grid_large <- dggridR::dgconstruct(area = params$hexagon_area_large)
# for an area of ~ 300 km^2, we get a resolution of 11:
grid_small <- dggridR::dgconstruct(area = params$hexagon_area_small)

```

```{r species-info}
ebirdst::ebirdst_runs %>% filter(substr(species_code,1,6) %in% species_codes$six)

```

```{r import-checklists}
checklists_path <- paste0(params$output_path, "/checklists.RDS")
filtered_checklists_path <- paste0(params$output_path, "/filtered_checklists.RDS")

if(params$always_import_checklists | !file.exists(checklists_path)){
  checklists <- import_checklists(params$erd_path)
  saveRDS(checklists, checklists_path)
}

if(params$always_filter_checklists | !file.exists(filtered_checklists_path)){
  checklists <- readRDS(checklists_path) %>%
    filter(latitude > params$extent_space$min_lat &
             latitude < params$extent_space$max_lat &
             longitude > params$extent_space$min_lon &
             longitude < params$extent_space$max_lon &
             year >= min(params$years) &
             year <= max(params$years) &
             ELEV_30M_MEDIAN < params$max_altitude &
             ((ELEV_30M_MEDIAN < params$max_altitude_above_lat42) | (latitude < 42)) &
             cci > params$effort_thresholds$cci_min &
             effort_distance_km <= params$effort_thresholds$dist_max &
             effort_hrs >= params$effort_thresholds$time_min &
             effort_hrs <= params$effort_thresholds$time_max
    ) %>%
    mutate(
      seqnum_large = dggridR::dgGEO_to_SEQNUM(
        grid_large, 
        .$longitude,
        .$latitude
        )[[1]],
      seqnum_small = dggridR::dgGEO_to_SEQNUM(
        grid_small, 
        .$longitude,
        .$latitude
        )[[1]]
      )
  saveRDS(checklists, filtered_checklists_path)
} else {
  checklists <- readRDS(filtered_checklists_path)
}

# extract unique large hexagons
cells_all <- unique(checklists$seqnum_large)

```

```{r bootstrap-abundance}
for(species_code in params$species_to_process){
  file_out <- paste0(params$output_path, "/abun_data/", species_code , ".rds")

  if(params$always_resample_bootstrap | !file.exists(file_out)){
    data <- 
      sample_grid_abun(
        species_code, params$erd_path, checklists, params$effort_thresholds, 
        params$extent_space, extent_time, time_window="full", 
        small_grid=grid_small$res, large_grid=grid_large$res, time_grid=7, 
        roi = raster_of_interest, quiet = FALSE
      )
    if(!dir.exists(dirname(file_out))) dir.create(dirname(file_out), recursive = TRUE)
    saveRDS(data, file_out)
  }
}

```

# Calculate spring/fall log-ratios 

```{r calculate-ratios-from-seasonal-abundances}
for(species_code in params$species_to_process){
  ##### load abundance data #####
  file_species <- paste0(params$output_path, "/abun_data/", species_code , ".rds")
  
  print(paste("loading data from file", file_species,"..."))
  data <- readRDS(file_species)
  
  ##### Get demographic indices #####
  cell_ratios <- get_ratios(data$abun, cells_all, n_small_min = params$n_small_min, quiet=params$quiet)
  tidy_ratios <- make_ratios_tidy(cell_ratios)
    
  # rename fall/spring ratios to productivity/recruitment:
  tidy_ratios$summary %>%
    mutate(season=ifelse(period=="fall", "prod","surv")) -> tidy_ratios$summary
  
  # save the ratio data
  saveRDS(tidy_ratios, paste0(params$output_path, "/abun_data/", species_code, "_ratios.rds"))
}
```

# plot demographic indices for specific cells
```{r plot ratio data for specific cells}
species_code="carwre"
tidy_ratios <- readRDS(paste0(params$output_path, "/abun_data/", species_code, "_ratios.rds"))

# select cells with at least 20 valid ratios
tidy_ratios$summary %>%
  group_by(cell) %>%
  summarize(sufficient_data = sum(is.na(median))<20) %>%  
  filter(sufficient_data) %>% pull(cell) -> cells_select

# plot selected cell 20 (example)
plot_ratios(cells_select[20], data=tidy_ratios$summary)
```


# compare recruitment/mortality variances

```{r compare-variance, echo=FALSE}
for(species_code in params$species_to_process){
  tidy_ratios <- readRDS(paste0(params$output_path, "/abun_data/", species_code, "_ratios.rds"))
  
  # compare variances in productivity and survival for each cell across years:
  var_save_path <- paste0(params$output_path, "/variance_results/", species_code,"/variance_test.rds")
  dir.create(dirname(var_save_path), recursive = TRUE, showWarnings = FALSE)
  if(params$always_run_variance_test | !file.exists(var_save_path)){
      print(paste0("processing_species ", species_code))
    data_compare_ratios <- lapply(sort(unique(tidy_ratios$summary$cell)),
                                  compare_ratio_variances, data=tidy_ratios$summary, 
                                  quiet=params$quiet) %>% do.call(rbind, .)
    saveRDS(data_compare_ratios,file=var_save_path)
  } else{
    data_compare_ratios <- readRDS(file=var_save_path)
  }
}
```

# compare bayesian/frequentist averages
Quick comparisons of inverse-sd weighted averages and the Bayesian estimates.
Using the last generated `tidy_ratios` object for now.
```{r compare-bayesian-frequentist averages}
# average indices over cells across years (using inverse weighting by sd)
# (note that n_prod,n_surv and n are identical within the grouping, so
# the weighted.mean just repeats the group value)
tidy_ratios$summary %>%
  group_by(cell,season) %>%
  filter(is.finite(avg)) %>%
  summarise(across(c("median","avg","q10","q90","sd","n_prod","n_surv","n", "has_inf"), \(x) weighted.mean(x, 1/sd, na.rm=TRUE)), .groups="drop_last") %>%
  mutate(has_inf=as.logical(has_inf)) -> data_cell
  
# join variance test results to year-averaged cell data:
data_cell <- left_join(data_cell,data_compare_ratios,by="cell")
```

# Download or load weather data
Download weather data for all relevant cells, if necessary.
```{r weather-data}
dir.create(paste0(params$output_path, "/weather"), showWarnings = FALSE)
weather_file <- paste0(params$output_path, "/weather/",paste0(params$daymet$label, collapse = "-"),".rds")

if(!params$quiet) print(paste("loading/processing weather file",basename(weather_file)))

# Warning messages below are from `erdPackage::daymet_extract` which calls
# `sp::getSpPPolygonsIDSlots`, which raises the deprecation warning
if(params$always_download_weather | !file.exists(weather_file)){
  # initialize google earth engine
  ee_Initialize()
  
  # helper function to extract daymet parameters specified by params$daymet
  extract_daymet_data <- function(year, cell, grid, params_daymet){
    # initialize the data.frame for daymet data:
    daymet_data=data.frame(matrix(nrow=1,ncol=nrow(params_daymet)+2,dimnames=list(NULL,c("cell","year",params_daymet$label))))
    daymet_data$cell=cell
    daymet_data$year=year
    # fill data.frame with averaged daymet data:
    for(lab in params_daymet$label){
      # filter parameters to use for daymet extraction:
      par_daymet <- params_daymet %>% filter(label==lab)
      # construct min and max dates
      maxdate=paste0(year,"-",par_daymet$date_max)
      mindate=paste0(year,"-",par_daymet$date_min)
      # address cases in which the min to max period includes a year change:
      if(as.Date(maxdate)<as.Date(mindate)) mindate=paste0(year-1,"-",par_daymet$date_min)
      # average over large cell, and store in data.frame
      daymet_data[par_daymet$label] <- daymet_extract(cell=cell, res=grid$res, variable=par_daymet$variable, mindate=mindate, maxdate=maxdate)
    }
    return(daymet_data)
  }
  
  # loop over cells and years
  data_daymet=data.frame()
  for (cell in cells_all) {
    for (year in params$years){
      data_daymet <- rbind(data_daymet, extract_daymet_data(year, cell, grid_large, params$daymet))
    }
  }
  saveRDS(data_daymet, weather_file)
} else{
  data_daymet <- readRDS(weather_file)
}
```
